---
title: "statComp_hw4"
author: "凌浩东"
date: "10/28/2021"
documentclass: ctexart
output: rticles::ctex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

For $x \in [a_{i-1},a_i]$, $i \in {1,\cdots,j}$,
$$
\begin{aligned}
P(X^*\leq x|J = i)& = \int_{-\infty}^{x} f_i(x)dx\\
&= k\int_{-a_{i-1}}^{x} f(x)dx\\
& = k(F(x) - \frac{i-1}{k})\\
& = kF(x) - i +1
\end{aligned}
$$
Hence,
$$
\begin{aligned}
P(X^*\leq x|J = i) = \left\{
\begin{array}{lcl}
kF(x) - i +1& & x \in [a_{i-1},a_i]\\
0 & & x \in (-\infty, a_{i-1})\\
1 & & x \in (a_i, \infty)
\end{array}
\right.
\end{aligned}
$$
Hence,
$$
\begin{aligned}
P(X^* \leq x) &=  \sum_{j = 1}^n P(X^*\leq x|J = j)P(J=j) \\
&= \frac{1}{k} \sum_{j = 1}^nP(X^*\leq x|J = j)\\
&= \frac{1}{k} \left(\sum_{j = 1}^{i-1}1 + P(X^*\leq x|J = i) + \sum_{j = i+1}^{n}0\right) \\
& = \frac{1}{k} \left(i-1 + [kF(x) - i +1]\right)\\
& = F(x)
\end{aligned}
$$
Hence, $X^* \sim X$.

For $Y^*$,
$$
Y^* = \frac{g_j(X^*)}{f_j(X^*)} = \frac{g_j(X^*)}{kf(X^*)} \sim \frac{g(X)}{kf(X)} = \frac{Y}{k}
$$
with $X^*\sim X$.


## 2

### exercise 6.1
```{r}
n <- 20
m <- 1000
mse <- numeric(9)
for (k in 1:9) {
  tmean <- numeric(m)
  for (i in 1:m) {
    x <- sort((rcauchy(n)))
    tmean[i] <- sum(x[(k+1):(n-k)])/(n-2*k)
  }
  target <- median(tmean)
  mse[k] <- mean((tmean - target)^2)
}
k <- c(1:9)
rbind(k, mse)
```
### exercise 6.9
```{r}

n <- 20
m <- 1000
gini.ratio <- numeric(m)
for (j in 1:m) {
  x <- sort(rlnorm(n))
  mu <- mean(x)
  G <- 0
  for (i in 1:n) {
    G <- G + (2*i - n - 1) * x[i]
  }
  gini.ratio[j] <- G / (n^2 * mu)  
}
mean <- mean(gini.ratio)
median <- median(gini.ratio)
c(mean, median)
quantile(gini.ratio, probs = seq(.1, .9, by = .1))
hist(gini.ratio, probability = T, 
     main = 'Histogram of Gini ratio generated 
     from log normal distribution', 
     xlab = 'Gini ratio')

```
```{r}

n <- 20
m <- 1000
gini.ratio <- numeric(m)
for (j in 1:m) {
  x <- sort(runif(n))
  mu <- mean(x)
  G <- 0
  for (i in 1:n) {
    G <- G + (2*i - n - 1) * x[i]
  }
  gini.ratio[j] <- G / (n^2 * mu)  
}
mean <- mean(gini.ratio)
median <- median(gini.ratio)
c(mean, median)
quantile(gini.ratio, probs = seq(.1, .9, by = .1))
hist(gini.ratio, probability = T, 
     main = 'Histogram of Gini ratio generated 
     from uniform distribution', 
     xlab = 'Gini ratio')
```
```{r}

n <- 100
m <- 1000
gini.ratio <- numeric(m)
for (j in 1:m) {
  x <- sort(rbinom(n, size = 1, prob = .1))
  mu <- mean(x)
  G <- 0
  for (i in 1:n) {
    G <- G + (2*i - n - 1) * x[i]
  }
  gini.ratio[j] <- G / (n^2 * mu)  
}
mean <- mean(gini.ratio)
median <- median(gini.ratio)
c(mean, median)
quantile(gini.ratio, probs = seq(.1, .9, by = .1))
hist(gini.ratio, probability = T, 
     main = 'Histogram of Gini ratio generated 
     from Bernoulli distribution', 
     xlab = 'Gini ratio')
```

## 3
As $X\sim N(0,1)$,
denote the pdf and cdf of standard normal as $f(x)$ and $F(x)$ respectively,
then

$$
\begin{aligned}
f_{X_{(i)}}(x) = \frac{n!}{(i-1)!(n-i)!}F^{i-1}(x)(1-F(x))^{n-i}f(x)\\
f_{X_{(n-i+1)}}(x) = \frac{n!}{(i-1)!(n-i)!}F^{n-i}(x)(1-F(x))^{i-1}f(x)\\
\end{aligned}
$$
Known $F(-x) = 1 - F(x)$,
$f(x) = f(-x)$,
we know that 
$$
\begin{aligned}
E\left[X_{(i)}+X_{(n-i+1)}\right]  =& \int_{-\infty}^{\infty} \frac{n!}{(i-1)!(n-i)!}F^{i-1}(x)(1-F(x))^{n-i}f(x)dx +\\ &\int_{-\infty}^{\infty}\frac{n!}{(i-1)!(n-i)!}F^{n-i}(x)(1-F(x))^{i-1}f(x)dx\\
= & \frac{n!}{(i-1)!(n-i)!}\left(\int_{-\infty}^{\infty} F^{i-1}(x)(1-F(x))^{n-i}f(x)dx +\right.\\ &\left.\int_{-\infty}^{\infty}F^{n-i}(-x)(1-F(-x))^{i-1}f(-x)dx\right)\\
= & \frac{n!}{(i-1)!(n-i)!}\left(\int_{-\infty}^{\infty} F^{i-1}(x)(1-F(x))^{n-i}f(x)dx +\right.\\ &\left.\int_{-\infty}^{\infty}(1-F(x))^{n-i}F^{i-1}(x)f(x)dx\right)\\
= & 0
\end{aligned}
$$
Hence,
$$
\begin{aligned}
E\left[\bar{X}_{[-k]}\right] & = E\left[\frac{1}{n-2k}\sum_{i=k+1}^{n-k}X_{(i)}\right]\\
& = \frac{1}{n-2k}\sum_{i=k+1}^{n-k}E\left[X_{(i)}\right]\\
& = \frac{1}{2(n-2k)}\sum_{i=k+1}^{n-k}E\left[X_{(i)}+X_{(n-i+1)}\right]\\
& = 0
\end{aligned}
$$

## 4

Denote $\frac{X_i-\mu}{\sigma}$ as $Z_i$,
$Y = AZ$, where $A$ is an  orthogonal matrix,
and the first row of $A$ is $(\frac{1}{\sqrt{n}}, \cdots, \frac{1}{\sqrt{n}})$,
then $Y \sim N(0, I)$,
$$
\lVert Y\rVert^2 = \sum_iY_i^2 \sim \chi^2(n)
$$
and
$$
Y_1 = \frac{\sum_i(X_i-\mu)}{\sqrt{n}\sigma} = \frac{\sqrt{n}(\bar{X} - \mu)}{\sigma} \sim N(0,1)
$$
where $Y_1$ is the first value of vector $Y$. 

Hence
$$
\begin{aligned}
\frac{\sum_i(X_i-\bar{X})^2}{\sigma^2} &= \frac{\sum_i(X_i-\mu)^2}{\sigma^2} - \frac{n(\bar{X} - \mu)^2}{\sigma^2}\\
& = \lVert Z\rVert^2 - Y_1^2\\
& = \lVert Y\rVert^2 - Y_1^2\\
& \sim \chi^2(n-1)
\end{aligned}
$$

and $\frac{\sum_i(X_i-\bar{X})^2}{\sigma^2}$ is independent with $Y_1 = \frac{\sqrt{n}(\bar{X} - \mu)}{\sigma}$,
i.e., $\frac{\sum_i(X_i-\bar{X})^2}{\sigma^2}$ is independent with $\bar{X}$














