---
title: "statComp_hw2"
author: "凌浩东"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1

## 5.6

$$
\begin{align*}
Cov(e^U, e^{1-U}) &=  
E(e^Ue^{1-U}) - Ee^UEe^{1-U} = 
e - (e - 1)^2
= -0.2342106\\

Var(e^U+e^{1-U}) & = Var(e^U) + Var(e^{1-U}) + 2Cov(e^U, e^{1-U}) \\
&= 2Var(e^U) + 2Cov(e^U, e^{1-U}) \\ 
&= 2(\frac{e^2 - 1}{2} - (e - 1)^2) + 2(e - (e - 1)^2)\\
& = 0.01564999
\end{align*}
$$
using Anthithetic variate, 
$\hat{\theta} = \frac{1}{2}e^U + \frac{1}{2}e^{1-U}$.

Hence
$$
\begin{align*}
Var(\hat{\theta}) = \frac{1}{4}Var(e^U + e^{1-U}) = 0.0039113
\end{align*}
$$
while $Var(e^U) = \frac{e^2 - 1}{2} - (e - 1)^2 = 0.2420351$.

Hence, the percent of reduction is 
$$
\frac{0.2420351 - 0.0039125}{0.2420351} \times 100\% = 98.4\%
$$

## 5.6


```{r}
set.seed(42)

m <- 10000
u <- runif(m)
T1 <- exp(u)
T2 <- exp(u) / 2 + exp(1-u) / 2

c(mean(T1), mean(T2))
(var(T1) - var(T2)) / var(T1)
```
The result is consistent with the theoretical value 
computed in 5.6.

## 5.8

$$
\begin{align*}
Cov(U, 1-U) &= E[U(1-U)] - EUE(1-U) = \frac{1}{6} - \frac{1}{2}\times \frac{1}{2} = -\frac{1}{12}\\
\rho (aU, a(1-U)) &= \frac{Cov(aU, a(1-U))}{\sqrt{Var(aU)Var(a(1-U))}} \\
&= \frac{a^2 Cov(U, 1-U)}{\sqrt{a^2Var(U)\times a^2Var(1-U)}} = \frac{Cov(U, 1-U)}{Var(U)} = -1
\end{align*}
$$
Suppose $U \sim Beta(\alpha, \alpha)$,
thus
$$
\begin{align*}
EU &= \frac{\alpha}{2\alpha} = \frac{1}{2}\\
EU^2 &= \frac{\alpha + 1}{2\alpha + 1}EU = \frac{\alpha + 1}{2(2\alpha + 1)}\\
Var(U) &= \frac{\alpha^2}{(\alpha+\alpha)^2(\alpha+\alpha+1)} = \frac{1}{4(2\alpha+1)}\\
Cov(U, 1-U) &= E[U(1-U)] - EUE(1-U) = \frac{1}{2} - \frac{\alpha + 1}{2(2\alpha + 1)} - \left(\frac{1}{2}\right)^2 \\
& = \frac{1}{4} - \frac{\alpha + 1}{2(2\alpha + 1)}\\
& = -\frac{1}{4(2\alpha+1)}\\
\rho (aU, a(1-U))&= \frac{a^2 Cov(U, 1-U)}{\sqrt{a^2Var(U)\times a^2Var(1-U)}} = \frac{Cov(U, 1-U)}{Var(U)} = -1
\end{align*}
$$

## 5.10

```{r}
set.seed(42)

MC.Myfun <- function(R = 10000, antithetic = T) {
  u <- runif(R/2)
  if (!antithetic) v <- runif(R/2) else
    v <- 1-u
  u <- c(u, v)
  g <- exp(-u) / (1+u^2)
  mean(g)
}

m <- 1000
T1 <- T2 <- numeric(m)
for (i in 1:m) {
  T1[i] <- MC.Myfun(antithetic = F)
  T2[i] <- MC.Myfun()
} 

c(mean(T1), mean(T2))
c(var(T1), var(T2))
print((var(T1) - var(T2)) / var(T1))
```

The estimator is 0.5248056 using Monte Carlo integration with antithetic variables.
The approximate reduction in variance is 96.30%.

# 2

> Monte Carlo method can be used to approximate the fraction of a d-dimensional
hypersphere which lies in the inscribed d-dimensional hypercube. Simulate with
different dimensions $d = 2, 3, 4, \cdots , 10$. (Hint: use `apply` function.)\\
(1) Derive the formula for the EXACT values for the above problem for each d-dimension.\\
(2) Using the above formula, approximate the value of $\pi$. Find the number of points
needed to approximate $\pi$ to its 4-th digit for each dimension d. Set the random
seed with `set.seed(123)` at the beginning of your R code.


```{r}
MC.ballVolumn <- function(d, m) {
  # generate matrix of uniform random variables
  # each column represent a point in the hyper space
  data <- matrix(runif(d*m), d, m)
  
  # for each point, check if the distance to the origin
  # if the distance smaller than 1 
  # the point is in the hypersphere
  p <- apply(data, 2, function(x) sum(x^2)<=1)
  
  # calculate how many points are in the hypersphere
  fraction <- apply(matrix(p, 1), 1, function(x) {sum(x)/m} )
  
  fraction
}

m <- 10000
fraction <- numeric(9)
for (d in c(2:10)) {
  fraction[d - 1] <- MC.ballVolumn(d, m)
}
fraction
```
```{r}
set.seed(123)


d <- 10
m <- 10000
result <- matrix(NA, d, m)
result <- outer(2:nrow(result), 1:ncol(result), 
                FUN = function(r,c) (2^r * gamma(r/2+1) * MC.ballVolumn(r, c))^(2/c))
result <- apply(result, 1, function(x) length(x[x - pi < 1e-4]))
result

```









