---
title: "statComp_hw3"
author: "凌浩东"
date: "10/21/2021"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Exercises 5.12, and 5.14.

### 5.12

Suppose $l \leq \frac{g(x)}{f(x)} \leq u$,
then

$$
\begin{align*}
Var(\hat{\theta}_{IS}) &= Var\left(\frac{1}{m}\sum_{i=1}^{m}\frac{g(x_i)}{f(x_i)}\right)\\
&= \frac{1}{m^2}\sum_{i=1}^{m}Var\left(\frac{g(x_i)}{f(x_i)}\right)\\
&= \frac{1}{m^2}\sum_{i=1}^{m}\left(\int\frac{g^2(x_i)}{f(x_i)}dx_i - \theta^2\right)
\end{align*}
$$

where

$$
l\theta = l\int g(x_i)dx_i \leq \int\frac{g^2(x_i)}{f(x_i)}dx_i \leq u\int g(x_i)dx_i = u\theta  
$$

thus

$$
\frac{1}{m}(l\theta - \theta^2) \leq Var(\hat{\theta}_{IS}) \leq \frac{1}{m}(u\theta - \theta^2)
$$
i.e.,  $Var(\hat{\theta}_{IS})$ is bounded.

### 5.14
First, let's take a look at our target function
$$
\begin{align*}
&g(x) = \frac{x^2}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} & x>1
\end{align*}
$$

```{r, include=FALSE}
library(ggplot2)
library(ggpmisc)
```

```{r}
g <- function(x) {
  x^2 * exp(-x^2 / 2) / sqrt(2*pi) * (x > 1) 
}

x <- seq(1, 10, length = 10000)
y <- g(x)
df <- data.frame(x, y)
c(x[ggpmisc:::find_peaks(df$y)], y[ggpmisc:::find_peaks(df$y)])
```

```{r}
ggplot(df, aes(x = x, y = y)) + geom_line() + stat_peaks(col = 'red') +
  scale_x_continuous(breaks=seq(0, 10, 1))
```

We can see that after a sharp ascend until 1.4 and descend to nearly 0 at around 4.
Hence we use two target function,
one is a shifted norm density function,
the other is a cauchy density function.

```{r}
set.seed(42)
m <- 10000

theta.hat <- se <- numeric(2)

x <- rnorm(m, mean = 1.4) 
fg <- g(x) / (dnorm(x, mean = 1.4))
theta.hat[1] <- mean(fg)
se[1] <- sd(fg)

x <- rcauchy(m, location = 1.4, scale = 1)
fg <- g(x) / dcauchy(x, location = 1.4, scale = 1)
theta.hat[2] <- mean(fg)
se[2] <- sd(fg)

theta.hat
se
```

## 2

> Given two random variables X and Y , prove the law of total variance

$$
var(Y) = E\{var(Y|X)\} + var\{E(Y|X)\}
$$

> Be explicit at every step of your proof.

$$
\begin{align*}
E\{var(Y|X)\} &= E\{E(Y^2|X) - [E(Y|X)]^2\}
= E(Y^2) - E\{[E(Y|X)]^2\} \\
var\{E(Y|X)\} &= E\{[E(Y|X)]^2\} - \{E[E(Y|X)]\}^2 = E\{[E(Y|X)]^2\} - [E(Y)]^2\\
\end{align*}
$$
Hence,
$$
E\{var(Y|X)\} + var\{E(Y|X)\} = E(Y^2) - [E(Y)]^2 = var(Y)
$$

## 3

> Define $\theta = \int_Ag(x)dx$, 
where A is a bounded set and $g \in \mathcal{L}_2(A)$. 
Let f be an importance function which is a density function supported on the set A.

> (a) Describe the steps to obtain the importance sampling estimator $\hat{\theta}_n$, 
where n is the number of random samples generated during the process.

* generate n random variables from f.

* caculate the mean of $\frac{g(x)}{f(x)}$, assign it to $\hat{\theta}_n$.

> (b) Show that the Monte Carlo variance of $\hat{\theta}_n$ is
$var(\hat{\theta}_n) = \frac{1}{n}\left\{\int_A\frac{g^2(x)}{f(x)}dx-\theta^2\right\}$.

$$
\begin{align*}
var(\hat{\theta}_n) &= var\left(\frac{1}{n}\sum_{i=1}^{n}\frac{g(x_i)}{f(x_i)}\right) \\
&= \frac{1}{n^2} \sum_{i = 1}^{n} var\left(\frac{g(x_i)}{f(x_i)}\right) \\
&= \frac{1}{n^2} \sum_{i = 1}^{n}\left(E\left[\left(\frac{g(x_i)}{f(x_i)}\right)^2\right] - \left(E\left[\frac{g(x_i)}{f(x_i)}\right]\right)^2\right) \\ 
& = \frac{1}{n^2} \sum_{i = 1}^{n} \left(\int_A\frac{g^2(x_i)}{f^2(x_i)}f(x_i)dx_i - \int_A\frac{g(x_i)}{f(x_i)}f(x_i)dx_i\right)\\
& = \frac{1}{n} \left\{\int_A\frac{g^2(x)}{f(x)}dx-\theta^2\right\}
\end{align*}
$$

> (c) Show that the optimal importance function $f^*$
, i.e., the minimizer of $var(\hat{\theta}_n)$, is
$f*(x) = \frac{|g(x)|}{\int_A|g(x)|dx}$,
and derive the theoretical lower bound of $var(\hat{\theta}_n)$.

From Cauchy-Schwartz inequality, 
we know that 
$$
\int_A\frac{g^2(x)}{f(x)}dx = \int_A\frac{g^2(x)}{f(x)}dx\cdot 1
= \int_A\frac{g^2(x)}{f(x)}dx \int_Af(x)dx \geq \left(\int_Ag(x)dx\right)^2
$$
where the equality holds when $f(x) \propto |g(x)|$.

Hence the minimizer is $f*(x) = \frac{|g(x)|}{\int_A|g(x)|dx}$.

The theoretical lower bound is $\frac{1}{n} \left\{\left(\int_Ag(x)dx\right)^2-\theta^2\right\} = 0$.









